---
title: "Assignment 6 Task 3"
author: "Logan Babcock"
date: "November 24, 2021"
output: html_document
---
# Introduction
This report will compare total lengths in millimeters for female lizards of two species *Cnemidophorus uniparens*, the Desert Grassland lizard, and *Cnemidophorus tessalatus*, the Colorado Checkered lizard. The data will analyzed using a Cohen's D test `cohen.d()`, to find the effect size of the difference in means. This wil let us assess if there is a significant difference in total length difference in mean total length between female lizards for the two species. 

```{r setup, include=FALSE, messages = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE )

library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)
library(effsize)

# read in data

lizards <- read_csv(here("data", "lizards.csv"))
```

```{r}
# tidying data and selecting variables we want
lizards_tidy <- lizards %>% 
  select("spp", "sex", "total_length") %>% # selecting columns
  filter(spp %in% c("CNUN", "CNTE")) %>% # operator selects specific vlues from range)
  mutate(Species = case_when( # reclassifying number values to political party 
    spp == "CNUN" ~ "Desert Grassland",
    spp == "CNTE" ~ "Colorado Checkered")) %>% # renaming variables 
  filter(sex == "F") # selecting females only
```
### Visualizing Lizard Lengths
```{r}
# visualizing data lengths
ggplot(lizards_tidy, aes(x = Species, y = total_length,
               colour = Species,
               shape = Species)) + 
  geom_boxplot(outlier.shape = NA) + # adding a boxplot to show spread
  geom_jitter() + 
  labs(y = "Total Length (mm)", x = "Female Lizards") +
  theme(axis.text.x = element_text(size = 7, hjust = 0.5), # aesthetic adjustments
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.caption = element_text(hjust = 0.9))

```
<center>**Figure 1:** Lengths of female lizards, faceted by species. Length is recorded in millimeters. Data spread is caputured by box plotes showing the means, and range of data. </center>

### Summary Table and Statistics

<center>**Table 1:** Summary table showing the mean, standard deviation, median length, and count of female lizards by species. </center>
```{r}
# making summary table
length_full <- na.omit(lizards_tidy)

summary_lengths <- length_full %>% # making summary chart
  group_by(Species) %>% # grouping by sex
  summarize(
    mean = mean(total_length), # summarizing mean, stand deviation, median, and counting sample size
    sd = sd(total_length),
    median = median(total_length),
    sample_size = n()) %>% 
  rename( "Standard Deviation" = "sd",
         "Median Length (mm)" = "median", 
         "Count" = "sample_size",
         "Mean" = "mean")  # renaming columns

kable(summary_lengths, digits = 2, align = "c") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```
```{r include = FALSE}
# exploratory graphs
ggplot(length_full, aes(sample = total_length)) + # exploratory qq plot faceted by species
  geom_qq() +
  facet_wrap(~Species) # looking at qq plots of each species

ggplot(length_full, aes(sample = total_length)) + # exploratory qq plot with population considered together
  geom_qq()


# making vectors for species lengths

colorado_length <- length_full %>% 
  filter(Species == "Colorado Checkered") %>% # selecting colorado values
  pull(total_length)

desert_length <- length_full %>% 
  filter(Species == "Desert Grassland") %>% 
  pull(total_length)

#exploratory ttest
length_ttest <- t.test(colorado_length, desert_length)
tidy_ttest <- tidy(length_ttest)
```

```{r}
# cohens effect size
length_cohen <- cohen.d(colorado_length, desert_length) # effect size

```
The effect size is large, with a Cohen's D of (Cohen's *d* = `r length_cohen$estimate`). The difference between the means is significant (two-sample t-test, t = `r round(tidy_ttest$statistic)`, p < 0.01). The actual difference in means is `r (summary_lengths$Mean [1]) - (summary_lengths$Mean [2])`. The large effect size and T-Test result means that we would reject the hypothesis that both species populations mean length's are the same

**Data Citation:** Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 



