---
title: "Assignment 6 Task 4: Mono Lake Data Visualization"
author: "Logan Babcock"
output: html_document
---
## Visualization of Mono Lake water level data from 1840 - 2016

```{r setup, include=FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(paletteer)
library(janitor)
library(here)
library(scales)
library(grid)

# read in data
mono <- read_csv(here("data","mono_lake_annual_levels.csv"), skip = 5)

# cleaning up the data
mono_tidy <- clean_names(mono) %>% 
  rename(lake_level = lake_level_feet_above_sea_level)

```

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
# visualizing the data
ggplot(data = mono_tidy, aes(x = year, y = lake_level)) +
  geom_line(aes(color = lake_level, legend = NA)) + # making line graph with no legend, and color change
  scale_x_continuous("Year",c(1850:2015), breaks = breaks_extended(20)) + # editing scale label 
  theme(panel.grid.minor = element_blank()) + # no grid lines
  theme(legend.title = element_blank(), legend.position = "none",  # axis labels and title font changes
        plot.title = element_text(hjust = 0.5, family = "serif"), 
        axis.text.x = element_text(family = "serif"), 
        axis.text.y = element_text(family = "serif"), 
        axis.title.x = element_text(family = "serif"), 
        axis.title.y = element_text(family = "serif")) +
  labs(x = "Year", y = "Lake Level (Feet above Sea Level)", # adding axis titles
       title = "Change in Water Level of Mono Lake", family = "serif") +
  annotate("text", x = 1890, y = 6379,  # landbridge eleveation annotation
           label = "Landbridge Elevation", 
           family = "serif", fontface = "italic", 
           colour = "black", size = 3) +
  annotate("text", x = 1890, y = 6362,  # salinity 1 annotaion
           label = "Lake Salinity increases to 120 g/L",
           family = "serif", fontface = "italic", 
           colour = "black", size = 3) +
  annotate("text", x = 1890, y = 6353, # salinity 2 annotation
           label = "Lake Salinity increases to 150 g/L",
           family = "serif", fontface = "italic", 
           colour = "black", size = 3) +
  geom_hline( aes(yintercept = 6377), linetype = "dashed", size = 0.25) + # horizontal lines for annotation
  geom_hline( aes(yintercept = 6360), linetype = "dashed", size = 0.25) +
  geom_hline( aes(yintercept = 6350), linetype = "dashed", size = 0.25) +
  annotate("segment", x = 1950, xend = 1940, y = 6425, yend = 6418, # arrow for start of diversion 
           colour = "black", size = 0.25, 
           arrow = arrow(length = unit(.25,"cm"))) +
  annotate("text", x = 1975, y = 6427, # diversion start annotation
           label = "Diversion of Mono Lake for L.A. water supply begins",
           family = "serif", 
           colour = "black", size = 3) +
  annotate("rect", xmin = 1850, xmax = 2020, ymin = 6345, ymax = 6377, # shaded area for salinity and landbridge
  alpha = .2) +
  annotate("text", x = 1890, y = 6395,  # mandated water level annotation
           label = "Mandated Water Level ( 6392 ft )",
           family = "serif", fontface = "italic", 
           colour = "black", size = 3) + 
  annotate("text", x = 1985, y = 6388, label = "Water Board enacts 
           water level mandate",
           family = "serif",fontface = "italic", 
colour = "black", size = 3) + # end of water board annotation
 annotate("segment", x = 1988, xend = 1994, y = 6384, yend = 6378,
           colour = "black", size = 0.25, 
          arrow = arrow(length = unit(.25,"cm"))) + # water board arrow
  geom_hline( aes(yintercept = 6392), linetype = "dashed", size = 0.25) # water mandadted level line
  
  

```
<center> **Figure 1:** Change in water level of Mono Lake, monitored from 1850 - 2016, measured in feet above sea level. Grey shaded area denotes the water level that reveals landbridges under the surface, allowing predation to occur and changing the perimeter of the lake. The salinity levels included mark the levels that would significantly detriment the availibilty of brine shrimp that are a critial food source of birds that Mono Lake provides critial habitat for. The State Water Board's decision in 1994 (Known as Decision 1631) is a cornerstone for Mono Lake conservation, and is the water level that the lake management is now aiming to reach. This level was chosen for its ecosystem services including ideal salinity level, and includes a buffer for anticipated drought years.  </center>

** Data Citation: ** “Mono Lake Levels 1850-Present.” Mono Lake Levels 1850-Present (Yearly), https://www.monobasinresearch.org/data/levelyearly.php.
