---
title: "Assignment 6 Task 2"
author: "Logan Babcock"
output: html_document
---

# Introduction

This report considers the survey data from the Environmental Data Initiative, which allows respondents to self-select political affiliation, and status as an environmentalist. The respondents can self select their party preference using a sliding scale, with 1-3 being conservative, and 5-7 being liberal. The polar ends represent stronger party affiliations. This report will select respondents within the 1-3 and 5-7 ranges, and run a Chi-Squared test `chisq.test()` to see if their response is related to their identification as an environmentalist. 


```{r setup, include=FALSE, echo = FALSE, messages = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE)

# read in packages 

library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)

# read in data 

survey <- read_csv(here("data", "conservation_survey.csv"))

```


```{r}
# filtering for the variables we want: POL_OR and ENV_ID
survey_tidy <- survey %>% 
  select(POL_OR, ENV_ID) %>% 
  filter(POL_OR %in% c("1","2","3","5","6","7")) %>% # operator selects specific vlues from range)
  mutate(Party = case_when( # reclassifying number values to political party 
    POL_OR == "1" ~ "Conservative",
    POL_OR == "2" ~ "Conservative",
    POL_OR == "3" ~ "Conservative",
    POL_OR == "5" ~ "Liberal",
    POL_OR == "6" ~ "Liberal",
    POL_OR == "7" ~ "Liberal")) %>% 
  mutate(env = case_when( # reclassifying env_id to anser yes, no, unsure
    ENV_ID == "1" ~ "Yes",
    ENV_ID == "2" ~ "No",
    ENV_ID == "3" ~ "Unsure"))


```
<center>**Table 1:** Porportions of responses for "Environmentalist" self-identificaition by political affiliation. "Yes" answers indicate survey respondent is an environmentalist, "No" indicates not an environmentalist. Numbers in parentheses are the number of responses for that category, while the percetage indicates the proportion of responses for that party affiliation.   </center>
```{r}
# making table showing proportions of reponses
tidy_counts <- survey_tidy %>% 
  tabyl(Party, env)

# getting proportions and counts together
proportions <- tidy_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% # updated sig figs
  janitor::adorn_ns() # includes count

# finalizing table 
kable(proportions, digits = 2, align = "c") %>% 
  kable_styling(bootstrap_options = (c("borered", "striped", "hover", full_width = F)))
```


```{r}
# contingency table
# moving party to a column
survey_ct <- tidy_counts %>%
  column_to_rownames(var = "Party") # separating the partyvariable into rows, not cell values

# chi squared test
survey_x2 <- chisq.test(survey_ct)


# cleaing up output
test_tidy <- tidy(survey_x2)
```
There is a signifiact assocation between party affiliation and if a person considered themselves an environmentalist, ($chi$^2^(`r round(test_tidy$parameter)`) = `r round(test_tidy$statistic)`, *p* = `r round(test_tidy$p.value)`). The proportions show that `r round(tidy_counts$No [1])` respondents that identified as Conservative marked "No", and only `r round(tidy_counts$No [2])` respondents that identified as Liberal marked "No". Since the Chi-Squared value is large, we can reject the null hypothesis that political affiliation and environmentalist idenfication are not associated. 

**Data citation:** Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c
