---
title: "Assignment 6 Task 3"
author: "Logan Babcock"
date: "November 24, 2021"
output: html_document
---
# Introduction
This report will compare total lengths in millimeters for female lizards of two species *Cnemidophorus uniparens*, the Desert Grassland lizard, and *Cnemidophorus tessalatus*, the Colorado Checkered lizard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)

# read in data

lizards <- read_csv(here("data", "lizards.csv"))
```

```{r}
# tidying data and selecting variables we want
lizards_tidy <- lizards %>% 
  select("spp", "sex", "total_length") %>% # selecting columns
  filter(spp %in% c("CNUN", "CNTE")) %>% # operator selects specific vlues from range)
  mutate(Species = case_when( # reclassifying number values to political party 
    spp == "CNUN" ~ "Desert Grassland",
    spp == "CNTE" ~ "Colorado Checkered")) %>% # renaming variables 
  filter(sex == "F") # selecting females only
```

```{r}
# visualizing data lengths
ggplot(lizards_tidy, aes(x = Species, y = total_length,
               colour = Species,
               shape = Species)) + 
  geom_boxplot(outlier.shape = NA) + # adding a boxplot to show spread
  geom_jitter() + 
  labs(y = "Total Length (mm)", x = "Female Lizards") +
  theme(axis.text.x = element_text(size = 7, hjust = 0.5), # aesthetic adjustments
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.caption = element_text(hjust = 0.9))

```
<center>**Figure 1:** Lengths of female lizards, faceted by species. Length is recorded in millimeters. Data spread is caputured by box plotes showing the means, and range of data. </center>
